---
title: "Site Selection Landcover Analyzer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sitetool}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Overview
The Site Selection Landcover Analyzer is an R package which launches an interactive R Shiny interface to assist with analyzing and selecting sites for field studies. The package also contains functions for analyzing spatial data outside of the interactive environment. 


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Interactive Version

The main component of the package is an Shiny interface that goes through the procedure of site selection step-by-step. To launch the interactive tool, execute the following commands:

```{r setup, eval=FALSE}
library(sitetool)
run_app()
```

This will open an interactive window. Follow the instructions in the "about" tab.

## Non-Interactive Version

The procedure can also be followed outside of the Shiny interface. These functions can be used on their own, or within a larger site selection procedure. 

### Get a list of sites

The package has two options for obtaining a list of potential sites: random sites or populated sites. Random sites are generated simply from a simple random sample within the study area (a provided shapefile or bounding box) and populated sites are any cities within the study area, obtained via OpenStreetMap.

The user must specify the area of interest (either via a bounding box or shapefile) and the number of desired points. To get a list of random sites, execute the following command:

```{r getlist, eval=FALSE, message=FALSE, warning=FALSE}
library(sitetool)
bbox <- c(-85, 29, -82, 31) #xmin, ymin, xmax, ymax
points <- get_random_points(area = bbox, n_points = 10)
```

The same command is also works with a shapefile, and the user can specify a distance from main roads and cities that points need to be located futher than:

```{r getlist2, eval=FALSE}
polygon <- st_read('path/to/shapefile.sf')
# road distance in meters, and city distance in kilometers
points <- get_random_points(area = polygon, n_points = 10, road_dist = 1000, city_dist = 10) 
```

Use the following command, to obtain a list of cities within the area of interest. The area can again be either a bounding box, or Polygon as an sf object. 

```{r getlistvillage,eval=FALSE}
bbox <- c(-85, 29, -82, 31) 
points <- get_village_points(area = bbox)
```


### Analyze landcover for a list of sites

Once you have a list of sites, you can analyze raster data surrounding each of the sites. The function takes in a raster, an sf object with a list of points, and the distance at which to calculate the landcover values. There are two options, one for categorical rasters (land-use/land-cover) and one for continuous rasters (NDVI, climate etc.).

You have two options for the landcover data. You can upload your own data:

```{r loadrast, eval=FALSE}
r = rast('path/to/rast.tif')
```

Or you can download data for an area of interest (as an sf object) using the following command: 

```{r dlrast, eval=FALSE}
r = get_cov_landuse(area)
```

Once you have a list of sites and a raster to analyze, execute the following command to create a dataframe:

```{r getlc, eval=FALSE}
df = createLCDataFrame(in_df = points, r=r, dist = 1000)
```

Or, if you have a continuous raster, exectue the following command:

```{r getndvi, eval=FALSE}
df = createNDVIDataFrame(in_df = points, r=r, dist = 1000)
```



## Development and issues
The \textit{sitetool} package is also available in a development version on GitHub: github.com/biodivhealth/ss-analyzer. Please log any issues or feature requests on GitHub. 
